---
title: "Dead coral removal_V2"
author: "Kai Kopecky"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Packages}
library(tidyverse)
library(janitor)
library(car)
library(lme4)
library(ggpattern)
library(chisq.posthoc.test)
library(DHARMa)
library(glmmTMB)
library(ggeffects)
library(calecopal)
library(emmeans)
library(glmmTMB)
```

```{r Globally used data}

# Live and dead coral annotations
regions_master <- read_csv("Data/Regions_master.csv") %>% 
  clean_names

```

## Time series
#### Proportion of live coral remaining over time
```{r Live coral data wrangling}
## Live coral raw data
# Create dataframe for live coral only
live_coral <- regions_master %>% 
  filter(tag_lab_class_name == "Pocillopora" | tag_lab_class_name == "Acropora") %>% 
  group_by(plot, treatment, tag_lab_date) %>% 
  summarize(live_coral.total_surf = sum(tag_lab_surf_area)) %>% 
  mutate(m_sq = live_coral.total_surf*0.0001,
         years_since_start = case_when(tag_lab_date == "8/1/19" ~ 0,
                                       tag_lab_date == "8/1/20" ~ 1,
                                       tag_lab_date == "8/1/21" ~ 2,
                                       tag_lab_date == "8/1/22" ~ 3,
                                       tag_lab_date == "8/1/23" ~ 4),
         time_point = case_when(tag_lab_date == "8/1/19" ~ 2019,
                                       tag_lab_date == "8/1/20" ~ 2020,
                                       tag_lab_date == "8/1/21" ~ 2021,
                                       tag_lab_date == "8/1/22" ~ 2022,
                                       tag_lab_date == "8/1/23" ~ 2023),
         time_point = as.factor(time_point))

## Proportional changes
# Create dataframe of initial amount of dead coral
live_coral_initial <- live_coral %>% 
  filter(years_since_start == 0) %>% 
  ungroup() %>% 
  rename(live_coral.initial = live_coral.total_surf) %>% 
  select(c(1,4))

# Create dataframe of proportional loss of live coral at each time point in each treatment
live_coral.prop <- merge(live_coral, live_coral_initial)
live_coral.prop <- live_coral.prop %>% 
  mutate(live_coral.prop_rem = live_coral.total_surf/live_coral.initial)

```

```{r Stats for live coral time series}

# Create dataframe without initial time point
live_coral.prop.shortened <- live_coral.prop %>% 
  filter(years_since_start != 0)

# Generalized linear mixed effects model of proportion of live coral remaining ~ the interaction of treatment and time point (categorical)
# Beta distribution used because response variable is proportional
# Plot identity as random effect to account for random variation among replicate plots and repeated measures of plots through time

live_coral.glmm.beta <- glmmTMB(live_coral.prop_rem ~ treatment*time_point + (1|plot),
               data = live_coral.prop.shortened,
               family = beta_family(link = "logit"))

summary(live_coral.glmm.beta)# Summary of GLMM 
hist(residuals(live_coral.glmm.beta), breaks = 20) # Histogram of residuals

# Generate predictions and confidence intervals from model, then turn into data frame
predictions.live_coral <- ggpredict(live_coral.glmm.beta, terms = ~time_point + treatment)
plot(predictions.live_coral)
predictions.live_coral <- as.data.frame(predictions.live_coral)

#Gives model predictions on scale of link function
emmeans.coral <- emmeans(live_coral.glmm.beta, ~ treatment | time_point) 

# Adjusted p-values for pairwise treatment comparisons in each year 
pairs(emmeans.coral, adjust = "bonferronni", type = "response")

```

```{r Visualization of live coral remaining over time}

# Add values for first time point back into dataframe
predictions.live_coral <- predictions.live_coral %>% 
  add_row(x = "2019", predicted = 1, std.error = 0, conf.low = 0, conf.high = 0, group = "Removal") %>%
  add_row(x = "2019", predicted = 1, std.error = 0, conf.low = 0, conf.high = 0, group = "Retention")

# Time series using model-predicted means and confidence intervals
live_coral.prop %>%
  ggplot(aes(x = time_point, y = live_coral.prop_rem))+
  geom_point(aes(shape = treatment),
             alpha = 0.5)+
  geom_line(data = predictions.live_coral, 
            aes(x = x , y = predicted, group = group))+
  geom_errorbar(data = predictions.live_coral, 
                aes(x = x, y = predicted, 
                    ymin =conf.low, ymax = conf.high),
                width = 0)+
  geom_point(data = predictions.live_coral, 
             aes(x = x, y = predicted, shape = group),
             size = 3)+
  geom_point(data = predictions.live_coral, 
             aes(x = x, y = predicted, shape = group),
             size = 2,
             color = "#CACFD0") +
  scale_y_continuous(limits = c(0, 1.1),
                     expand = c(0, 0)) +
  labs(x = "Year",
       y = "Proportion of live coral remaining") +
  scale_shape_manual(values = c(19,17),
                     labels = c("Removal", "Retention"),
                     name = "Skeleton treatment") +
  #coord_fixed(ratio = 3) +
  theme_classic(base_size = 14)

```

```{r Live coral cover in 2023}
# Live coral cover in 2023
live_coral.2023 <- live_coral %>% 
  filter(time_point == "2023") 

live_coral.2023.summary <- live_coral.2023 %>% 
  group_by(treatment) %>% 
  summarize(mean_cover = mean(m_sq),
            SE = sd(m_sq)/sqrt(n()))

ggplot(live_coral.2023.summary, aes(x = treatment, y = mean_cover, shape = treatment)) +
  geom_errorbar(aes(ymax = mean_cover + SE,
                    ymin = mean_cover - SE),
                width = 0) +
  geom_point(data = live_coral.2023, aes(x = treatment, y = m_sq),
             size = 2,
             alpha = 0.5) +
  geom_point(size = 4) +
  geom_point(size = 3,
             color = "#CACFD0") +
  labs(x = "Skeleton treatment",
       y = Surface~area~of~live~coral~(m^2)) +
  coord_fixed(ratio = 1.5) +
  scale_y_continuous(limits = c(0,2.2),
                     expand = c(0,0)) +
  scale_shape_manual(values = c(19,17),
                    name = "Skeleton treatment",
                    labels = c("Removal", "Retention")) +
  theme_classic(base_size = 12)

```


#### Percent cover of macroalgae over time
```{r Data wrangling for macroalgae cover}

# Load raw data and calculate percent cover
algae <- read_csv("Data/Algae point data.csv") %>% 
  clean_names() %>% 
  mutate(time_point = as.factor(time_point),
         prop_cover = (points/total_points),
         perc_cover = prop_cover*100)

# Create summary table for total algae cover in each treatment/time point 
algae_totals <- algae %>% 
  group_by(plot, treatment, time_point) %>% 
  summarize(total_cover.prop = sum(prop_cover),
            total_cover = sum(perc_cover))

```

```{r Stats for macroalgae time series}

# Create dataframe without initial time point
macroalgae.prop.shortened <- algae_totals %>% 
  filter(time_point != "2019")

# Generalized linear mixed effects model of proportional cover of macroalgae ~ the interaction of treatment and time point (categorical)
# Beta distribution used because response variable is proportional
# Plot identity as random effect to account for random variation among replicate plots and repeated measures of plots through time

algae.glmm.beta <- glmmTMB(total_cover.prop ~ treatment*time_point + (1|plot),
               data = macroalgae.prop.shortened,
               family = beta_family(link = "logit"))

summary(algae.glmm.beta)# Summary of GLMM 
hist(residuals(algae.glmm.beta), breaks = 20) # Histogram of residuals

# Generate predictions and confidence intervals from model, then turn into data frame
predictions.algae <- ggpredict(algae.glmm.beta, terms = ~time_point + treatment)
plot(predictions.algae)
predictions.algae <- as.data.frame(predictions.algae)

#Gives model predictions on scale of link function
emmeans.algae <- emmeans(algae.glmm.beta, ~ treatment | time_point) 

# Adjusted p-values for pairwise treatment comparisons in each year 
pairs(emmeans.algae, adjust = "bonferronni", type = "response")

```

```{r Visualization for algae time series}

# Add values for first time point back into dataframe
predictions.algae <- predictions.algae %>% 
  add_row(x = "2019", predicted = 0.00229, std.error = 0.001, conf.low = 0.00201, conf.high = 0.00201, group = "Removal") %>%
  add_row(x = "2019", predicted = 0.0109, std.error = 0.00232, conf.low = 0.00464, conf.high = 0.00464, group = "Retention")

# Time series using model-predicted means and confidence intervals (multiplied by 100 to get percent cover)
algae_totals %>%
  ggplot(aes(x = time_point, y = total_cover))+
  geom_point(aes(shape = treatment),
             alpha = 0.5) +
  geom_line(data = predictions.algae, 
            aes(x = x , y = predicted*100, group = group))+
  geom_errorbar(data = predictions.algae, 
                aes(x = x, y = predicted*100, 
                    ymin =conf.low*100, ymax = conf.high*100),
                width = 0)+
  geom_point(data = predictions.algae, 
             aes(x = x, y = predicted*100, shape = group),
             size = 3)+
  geom_point(data = predictions.algae, 
             aes(x = x, y = predicted*100, shape = group),
             size = 2,
             color = "#CACFD0") +
  scale_y_continuous(limits = c(0, 40),
                     expand = c(0, 0)) +
  labs(x = "Year",
       y = "Percent cover of macroalgae") +
  scale_shape_manual(values = c(19,17),
                     labels = c("Removal", "Retention"),
                     name = "Skeletal treatment") +
  #coord_fixed(ratio = 3) +
  theme_classic(base_size = 14)

```

```{r Algae cover in 2023}

algae.2023 <- algae_totals %>% 
  filter(time_point == "2023") 

algae.2023.summary <- algae.2023 %>% 
  group_by(treatment) %>% 
  summarize(mean_cover = mean(total_cover),
            SE = sd(total_cover)/sqrt(n()))

ggplot(algae.2023.summary, aes(x = treatment, y = mean_cover, shape = treatment)) +
  geom_errorbar(aes(ymax = mean_cover + SE,
                    ymin = mean_cover - SE),
                width = 0) +
  geom_point(data = algae.2023, aes(x = treatment, y = total_cover),
             size = 2,
             alpha = 0.5) +
  geom_point(size = 4) +
  geom_point(size = 3,
             color = "#CACFD0") +
  labs(x = "Skeleton treatment",
       y = "Percent cover of macroalgae") +
  coord_fixed(ratio = 0.15) +
  scale_y_continuous(limits = c(0,25),
                     expand = c(0,0)) +
  scale_shape_manual(values = c(19,17),
                    name = "Skeleton treatment",
                    labels = c("Removal", "Retention")) +
  theme_classic(base_size = 12)
```

#### Algae time series separated by algal taxa
```{r}

# Create summary table of algae by taxa
algae_taxa.summary <- algae %>% 
  mutate(time_point = as.numeric(time_point),
         time_point = case_when(time_point == 1 ~ 2019,
                                time_point == 2 ~ 2020,
                                time_point == 3 ~ 2021,
                                time_point == 4 ~ 2022,
                                time_point == 5 ~ 2023)) %>% 
  group_by(treatment, algal_taxa, time_point) %>% 
  summarize(mean_cover.taxa = mean(perc_cover)) %>% 
  mutate(algal_taxa = as.factor(algal_taxa))
 
algae_taxa.summary$algal_taxa <- factor(algae_taxa.summary$algal_taxa, levels = c("Lobophora", "Asparagopsis", "Halimeda", "Turbinaria"))

ggplot(algae_taxa.summary, aes(x = time_point, y = mean_cover.taxa, fill = algal_taxa)) +
  geom_area(linewidth = 0.5, 
            colour="black") +
  facet_wrap(~treatment) +
  labs(x = "Year",
       y = "Percent cover of macroalgae") +
  scale_fill_manual(values = c("#44781E", "#2C3B75", "#A1B654", "#565052"),
                    name = 'Algal taxa') +
  theme_classic(base_size = 14)

## Stacked area chart with proportions of algae rather than total percent cover, not separated by treatment
algae_prop <- algae %>%
  group_by(plot, algal_taxa, time_point) %>%
  summarize(taxa_cover = sum(perc_cover))

algae_prop <- merge(algae_prop, algae_totals)

algae_prop.summary <- algae_prop %>%
  mutate(time_point = as.numeric(time_point),
         time_point = case_when(time_point == 1 ~ 2019,
                                time_point == 2 ~ 2020,
                                time_point == 3 ~ 2021,
                                time_point == 4 ~ 2022,
                                time_point == 5 ~ 2023)) %>% 
  mutate(prop_cover = taxa_cover/total_cover) %>%
  group_by(algal_taxa, time_point) %>%
  summarize(mean_prop = mean(prop_cover)) %>%
  mutate(mean_prop = replace_na(mean_prop, 0))

algae_prop.summary$algal_taxa <- factor(algae_prop.summary$algal_taxa, levels = c("Lobophora", "Asparagopsis", "Halimeda", "Turbinaria"))

ggplot(algae_prop.summary, aes(x = time_point, y = mean_prop, fill = algal_taxa)) +
  geom_area(linewidth = 0.5,
            colour="black") +
  labs(x = "Year",
       y = "Proportion of macroalgal cover") +
  scale_fill_manual(values = c("#44781E", "#2C3B75", "#A1B654", "#565052"),
                    name = 'Macroalgal taxa',
                    labels = c(expression(italic("Lobophora")), 
                               expression(italic("Asparagopsis")),
                               expression(italic("Halimeda")), 
                               expression(italic("Turbinaria")))) +
  theme_classic(base_size = 14)


```

#### Algae associations to dead coral
```{r Data wrangling and pie charts}
# Create dataframe for proportions of points inside and outside of dead coral regions
algae_associations <- algae %>% 
  drop_na() %>% 
  group_by(algal_taxa) %>% 
  summarize(total_points = sum(points),
            total_points.in_dc = sum(points_within_dead_coral)) %>% 
  mutate(inside = total_points.in_dc/total_points,
         outside = 1 - inside) %>% 
  pivot_longer(cols = c(inside, outside),
               values_to = 'proportion',
               names_to = 'position')

# Pie charts
# Lobophora
lobophora <- algae_associations %>% 
  filter(algal_taxa == "Lobophora") %>% 
  select(position, proportion)

ggplot(lobophora, aes(x = "", y = proportion, fill = position)) +
  geom_col_pattern(width=1, 
                   color= "#CACFD0",
                   pattern_colour = NA,
                   pattern_fill = "#CACFD0",
                   aes(pattern = position)) +
  coord_polar("y", start=0) +
  scale_pattern_discrete(choices = c("stripe", "none")) +
  scale_fill_manual(values = c("#44781E", "#44781E"),
                    name = "",
                    labels = c("Inside", "Outside")) +
  #facet_wrap(~treatment) +
  theme_void()

# Asparagopsis
asparagopsis <- algae_associations %>% 
  filter(algal_taxa == "Asparagopsis") %>% 
  select(position, proportion)

ggplot(asparagopsis, aes(x = "", y = proportion, fill = position)) +
  geom_col_pattern(width=1, 
                   color= "#CACFD0",
                   pattern_colour = NA,
                   pattern_fill = "#CACFD0",
                   aes(pattern = position)) +
  coord_polar("y", start=0) +
  scale_pattern_discrete(choices = c("stripe", "none")) +
  scale_fill_manual(values = c("#2C3B75", "#2C3B75"),
                    name = "",
                    labels = c("Inside", "Outside")) +
  #facet_wrap(~treatment) +
  theme_void()

# Halimeda
halimeda <- algae_associations %>% 
  filter(algal_taxa == "Halimeda") %>% 
  select(position, proportion)

ggplot(halimeda, aes(x = "", y = proportion, fill = position)) +
  geom_col_pattern(width=1, 
                   color= "#CACFD0",
                   pattern_colour = NA,
                   pattern_fill = "#CACFD0",
                   aes(pattern = position)) +
  coord_polar("y", start=0) +
  scale_pattern_discrete(choices = c("stripe", "none")) +
  scale_fill_manual(values = c("#A1B654", "#A1B654"),
                    name = "",
                    labels = c("Inside", "Outside")) +
  #facet_wrap(~treatment) +
  theme_void()

# Turbinaria
turbinaria <- algae_associations %>% 
  filter(algal_taxa == "Turbinaria") %>% 
  select(position, proportion)

ggplot(turbinaria, aes(x = "", y = proportion, fill = position)) +
  geom_col_pattern(width=1, 
                   color= "#CACFD0",
                   pattern_colour = NA,
                   pattern_fill = "#CACFD0",
                   aes(pattern = position)) +
  coord_polar("y", start=0) +
  scale_pattern_discrete(choices = c("stripe", "none")) +
  scale_fill_manual(values = c("#565052", "#565052"),
                    name = "",
                    labels = c("Inside", "Outside")) +
  #facet_wrap(~treatment) +
  theme_void()

# to get blank legend
ggplot(asparagopsis, aes(x = "", y = proportion, fill = position)) +
  geom_col_pattern(width=1, 
                   color= "#CACFD0",
                   pattern_colour = NA,
                   pattern_fill = "#CACFD0",
                   aes(pattern = position)) +
  coord_polar("y", start=0) +
  scale_pattern_discrete(choices = c("stripe", "none")) +
  scale_fill_manual(values = c("black", "black"),
                    name = "Location of points",
                    labels = c("On dead skeletons", "On reef")) +
  theme_void()

```

```{r Stats for algae associations}

## Treatments combined
algae_contingency <- algae %>% 
  drop_na() %>%
  group_by(algal_taxa) %>% 
  summarize(total_points = sum(points),
            inside = sum(points_within_dead_coral),
            outside = total_points - inside) %>% 
  select(-total_points, -algal_taxa)

algae_matrix <- as.matrix(algae_contingency, ncol = 2, nrow = 4)
rownames(algae_matrix) <- c("Asparagopsis", "Halimeda", "Lobophora", "Turbinaria")
colnames(algae_matrix) <- c("Inside", "Outside")

chi_sq.algae <- chisq.test(algae_matrix)
chi_sq.algae$expected
chi_sq.algae$observed
chi_sq.algae.post_hoc <- chisq.posthoc.test(algae_matrix)


## Separated by treatment
# Contingency table for algae associations: removal plots
algae_contingency.removal <- algae %>% 
  drop_na() %>% 
  filter(treatment == "Removal") %>% 
  group_by(algal_taxa) %>% 
  summarize(total_points = sum(points),
            inside = sum(points_within_dead_coral),
            outside = total_points - inside) %>% 
  select(-total_points, -algal_taxa)

algae_matrix.removal <- as.matrix(algae_contingency.removal, ncol = 2, nrow = 4)
rownames(algae_matrix.removal) <- c("Asparagopsis", "Halimeda", "Lobophora", "Turbinaria")
colnames(algae_matrix.removal) <- c("Inside", "Outside")

chi_sq.algae.removal <- chisq.test(algae_matrix.removal)
chi_sq.algae.removal$expected
chi_sq.algae.removal$observed
chi_sq.algae..removal.post_hoc <- chisq.posthoc.test(algae_matrix.removal) # All p-values are zero... this seems to be real, just lots of data so departures from expected values are very strong and clear

# Contingency table for algae associations: retention plots
algae_contingency.retention <- algae %>% 
  drop_na() %>% 
  filter(treatment == "Retention") %>% 
  group_by(algal_taxa) %>% 
  summarize(total_points = sum(points),
            inside = sum(points_within_dead_coral),
            outside = total_points - inside) %>% 
  select(-total_points, -algal_taxa)

algae_matrix.retention <- as.matrix(algae_contingency.retention, ncol = 2, nrow = 4)
rownames(algae_matrix.retention) <- c("Asparagopsis", "Halimeda", "Lobophora", "Turbinaria")
colnames(algae_matrix.retention) <- c("Inside", "Outside")

chi_sq.algae.retention <- chisq.test(algae_matrix.retention)
chi_sq.algae.retention$expected
chi_sq.algae.retention$observed
chi_sq.algae..retention.post_hoc <- chisq.posthoc.test(algae_matrix.retention)


##### Algae associations, plot averages
algae_associations.plots <- algae %>% 
  drop_na() %>% 
  group_by(plot, treatment, algal_taxa) %>% 
  summarize(total_points = sum(points),
            total_points.in_dc = sum(points_within_dead_coral)) %>% 
  mutate(inside = total_points.in_dc/total_points,
         outside = 1 - inside) %>% 
  pivot_longer(cols = c(inside, outside),
               values_to = 'proportion',
               names_to = 'position') %>% 
  drop_na()

algae_associations.plots.summary <- algae_associations.plots %>% 
  group_by(treatment, algal_taxa, position) %>% 
  summarize(mean_prop = mean(proportion),
            se_prop = sd(proportion)/sqrt(n()))

ggplot(algae_associations.plots.summary, aes(x = algal_taxa, y = mean_prop, fill = position)) +
  geom_col(position = position_dodge(),
           color = "black") +
  geom_errorbar(aes(ymax = mean_prop + se_prop,
                    ymin = mean_prop - se_prop,
                    width = 0),
                width = 0.9,
                position = position_dodge()) +
  facet_wrap(~treatment) +
  theme_classic()

```

```{r}
# Mixed effects model of algae associations
algae_associations.glmm <- algae_associations %>% 
  select(-position, -proportion)

# Create the two-column response variable
algae_associations.glmm$response <- cbind(algae_associations.glmm$total_points.in_dc, algae_associations.glmm$total_points - algae_associations.glmm$total_points.in_dc)

# Fit the mixed-effects model
associations.glmm <- glmmTMB(response ~ algal_taxa*treatment, 
                 family = binomial(link = "logit"), 
                 data = algae_associations.glmm)

# Model summary
summary(associations.glmm)

```


## Relationships between dead coral, algae, and live coral 
#### Algae cover ~ dead coral cover
```{r Data wrangling for dead coral cover and algae}

# Create dataframe of dead coral cover
dead_coral <- regions_master %>%
  filter(tag_lab_class_name == "Dead coral") %>% 
  group_by(plot, treatment, tag_lab_date) %>% 
  summarize(dead_coral.m_sq = sum(tag_lab_surf_area)*0.0001) %>% 
  mutate(time_point = case_when(tag_lab_date == "8/1/19" ~ 2019,
                                       tag_lab_date == "8/1/20" ~ 2020,
                                       tag_lab_date == "8/1/21" ~ 2021,
                                       tag_lab_date == "8/1/22" ~ 2022,
                                       tag_lab_date == "8/1/23" ~ 2023),
         time_point = as.factor(time_point))

# Build merged dataframe of dead coral and macroalgae cover
algae.dead_coral <- merge(algae_totals, dead_coral)

```

```{r Statistics for algae cover ~ dead coral}

# Remove initial time point
algae.dead_coral.short <- algae.dead_coral %>% 
  filter(time_point != "2019")

# Generalized linear mixed effects model of proportional cover of macroalgae ~ dead coral cover, time point and plot as random effects
algae.dead_coral.glmm <- glmmTMB(total_cover.prop ~ dead_coral.m_sq + (1|time_point) + (1|plot), 
                    family = beta_family(link = "logit"),
                    data = algae.dead_coral.short)
summary(algae.dead_coral.glmm)

# Get confidence intervals for model
confint(algae.dead_coral.glmm)

# Explore residuals of model for normality
hist(residuals(algae.dead_coral.glmm))
car::qqPlot(residuals(algae.dead_coral.glmm))

# Plot of predicted relationship
plot(ggpredict(algae.dead_coral.glmm, terms = ~dead_coral.m_sq))

# Create dataframe of model predictions
algae.dead_coral.predictions <- as.data.frame(ggpredict(algae.dead_coral.glmm, terms = ~dead_coral.m_sq))

```

```{r Visualization of algae cover ~ dead coral cover}

algae.dead_coral.short %>%
  ggplot()+
  geom_ribbon(data = algae.dead_coral.predictions, 
              aes(x = x, y = predicted*100, ymin =conf.low*100, ymax = conf.high*100), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = algae.dead_coral.predictions, 
            aes(x = x , y = predicted*100))+
  geom_point(aes(x = dead_coral.m_sq, y = total_cover, shape = treatment),
             alpha = 0.6,
             size = 2)+
  labs(x = Dead~coral~cover~(m^2),
       y = "% cover of macroalgae") +
  scale_shape_manual(values = c(19,17),
                     labels = c("Removal", "Retention"),
                     name = "Skeleton treatment") +
  theme_classic(base_size = 14)

```

#### Annual loss of live coral ~ annual macroalgae cover
```{r Data wrangling for live coral loss ~ algae cover}

# Create dataframe for proportional changes in live coral in each year (time period)
coral_change.prop <- live_coral %>%
  select(-c(tag_lab_date, live_coral.total_surf, years_since_start)) %>%
  rename(live_coral.m_sq = m_sq) %>%
  pivot_wider(names_from = time_point, values_from = live_coral.m_sq) %>%
  rename(year_0 = "2019",
         year_1 = "2020",
         year_2 = "2021",
         year_3 = "2022",
         year_4 = "2023") %>%
  mutate(delta.1 = (year_1 - year_0)/year_0,
         delta.2 = (year_2 - year_1)/year_1,
         delta.3 = (year_3 - year_2)/year_2,
         delta.4 = (year_4 - year_3)/year_3,
         delta.1.raw = year_1 - year_0,
         delta.2.raw = year_2 - year_1,
         delta.3.raw = year_3 - year_2,
         delta.4.raw = year_4 - year_3)

# # Proportional coral change
# coral_change.prop <- coral_change.prop %>%
#   select(-c(year_0, year_1, year_2, year_3, year_4)) %>%
#   pivot_longer(cols = c("delta.1", "delta.2", "delta.3", "delta.4"),
#                names_to = "delta_year",
#                values_to = "delta_coral") %>%
#   drop_na() %>%
#   mutate(period = case_when(delta_year == "delta.1" ~ "2019 - 2020",
#                           delta_year == "delta.2" ~ "2020 - 2021",
#                           delta_year == "delta.3" ~ "2021 - 2022",
#                           delta_year == "delta.4" ~ "2022 - 2023"))

# Raw coral change
coral_change.raw <- coral_change.prop %>%
  select(-c(year_0, year_1, year_2, year_3, year_4)) %>%
  pivot_longer(cols = c("delta.1.raw", "delta.2.raw", "delta.3.raw", "delta.4.raw"),
               names_to = "delta_year",
               values_to = "delta_coral.raw") %>%
  drop_na() %>%
  mutate(period = case_when(delta_year == "delta.1.raw" ~ "2019 - 2020",
                          delta_year == "delta.2.raw" ~ "2020 - 2021",
                          delta_year == "delta.3.raw" ~ "2021 - 2022",
                          delta_year == "delta.4.raw" ~ "2022 - 2023"))

# # Create dataframe for average algae cover in each time period
# algae_means <- algae_totals %>%
#   select(-total_cover) %>% 
#   pivot_wider(names_from = time_point, values_from = total_cover.prop) %>%
#   rename(year_0 = "2019",
#          year_1 = "2020",
#          year_2 = "2021",
#          year_3 = "2022",
#          year_4 = "2023") %>%
#   mutate(mean.1 = (year_1 + year_0)/2,
#          mean.2 = (year_2 + year_1)/2,
#          mean.3 = (year_3 + year_2)/2,
#          mean.4 = (year_4 + year_3)/2)
# 
# algae_means <- algae_means %>%
#   select(-c(year_0, year_1, year_2, year_3, year_4)) %>%
#   pivot_longer(cols = c("mean.1", "mean.2", "mean.3", "mean.4"),
#                names_to = "period",
#                values_to = "algae.mean") %>%
#   drop_na() %>%
#   mutate(period = case_when(period == "mean.1" ~ "2019 - 2020",
#                           period == "mean.2" ~ "2020 - 2021",
#                           period == "mean.3" ~ "2021 - 2022",
#                           period == "mean.4" ~ "2022 - 2023"))

# algae_cover <- algae_means %>%
#   select(year_1, year_2, year_3, year_4) %>%
#   pivot_longer(cols = c("year_1", "year_2", "year_3", "year_4"),
#                names_to = "year",
#                values_to = "algae.cover") %>%
#   drop_na() %>%
#   mutate(year = case_when(year == "year_1" ~ "2020",
#                           year == "year_2" ~ "2021",
#                           year == "year_3" ~ "2022",
#                           year == "year_4" ~ "2023"))

# # Proportional coral change
# algae_means.coral_change <- merge(coral_change.prop, algae_means)
# algae_means.coral_change <- algae_means.coral_change %>% 
#   mutate(delta_coral.trans = delta_coral*(-1))
# 
# # Raw coral change
# algae_means.coral_change.raw <- merge(coral_change.raw, algae_means)
# algae_means.coral_change.raw <- algae_means.coral_change.raw %>% 
#   mutate(delta_coral.trans = delta_coral.raw*(-1))

# Raw coral change, raw algae cover
algae_totals.coral_change.raw <- algae_totals %>% 
  filter(time_point != "2019") %>% 
  mutate(period = case_when(time_point == "2020" ~ "2019 - 2020",
                          time_point == "2021" ~ "2020 - 2021",
                          time_point == "2022" ~ "2021 - 2022",
                          time_point == "2023" ~ "2022 - 2023")) %>% 
  select(-time_point, -total_cover)

coral_change.raw <- coral_change.raw %>% 
  select(-c(3:7))

algae_totals.coral_change.raw <- merge(coral_change.raw, algae_totals.coral_change.raw)
# algae_totals.coral_change.raw <- algae_means.coral_change.raw %>% 
#   mutate(delta_coral.trans = delta_coral.raw*(-1))

```

```{r Stats for change in coral ~ algae cover}

# # Linear mixed effects model of annual proportional change in coral cover ~ mean algae cover in the same period; time period and plot identity as random effects
# coral.algae.glmm <- glmmTMB(delta_coral ~ algae.mean + (1|period) + (1|plot), 
#                     family = gaussian(link = "identity"),
#                     data = algae_means.coral_change)
# summary(coral.algae.glmm)
# 
# # Get confidence interval for model
# confint(coral.algae.glmm)
# 
# # Check residuals
# hist(residuals(coral.algae.glmm))
# car::qqPlot(residuals(coral.algae.glmm))
# 
# # Plot model predictions and create dataframe 
# plot(ggpredict(coral.algae.glmm, terms = ~algae.mean))
# coral.algae.predictions <- as.data.frame(ggpredict(coral.algae.glmm, terms = ~algae.mean))

```

```{r Visualizaton of proportional coral change ~ mean algae cover in each period}

# # Algae multiplied by 100 to be expressed in percent cover
# algae_means.coral_change %>%
#   ggplot()+
#   geom_ribbon(data = coral.algae.predictions, 
#               aes(x = x*100, y = predicted, ymin =conf.low, ymax = conf.high),
#               alpha =0.5,
#               fill = "#3B4F8E")+
#   geom_line(data = coral.algae.predictions, 
#             aes(x = x*100, y = predicted))+
#   geom_point(aes(x = algae.mean*100, y = delta_coral, shape = treatment),
#              alpha = 0.6,
#              size = 2)+
#   labs(x = "Annual % cover of macroalgae",
#        y = "Proportional change in live coral") +
#   scale_shape_manual(values = c(19,17),
#                      labels = c("Removal", "Retention"),
#                      name = "Skeletal treatment") +
#   theme_classic(base_size = 14)
# 
# # Algae and coral multiplied by 100 to be expressed in percent cover
# algae_means.coral_change %>%
#   ggplot()+
#   geom_ribbon(data = coral.algae.predictions, 
#               aes(x = x*100, y = predicted*100, ymin =conf.low*100, ymax = conf.high*100), 
#               alpha =0.5,
#               fill = "#3B4F8E")+
#   geom_line(data = coral.algae.predictions, 
#             aes(x = x*100, y = predicted*100))+
#   geom_point(aes(x = algae.mean*100, y = delta_coral*100, shape = treatment),
#              alpha = 0.6,
#              size = 2)+
#   labs(x = "% cover of macroalgae",
#        y = "Annual % change in live coral") +
#   scale_shape_manual(values = c(19,17),
#                      labels = c("Removal", "Retention"),
#                      name = "Skeleton treatment") +
#   theme_classic(base_size = 14)
```

```{r Stats for raw change in coral ~ algae cover}
# 
# # Linear mixed effects model of annual proportional change in coral cover ~ mean algae cover in the same period; time period and plot identity as random effects
# coral.algae.glmm.raw <- glmmTMB(delta_coral.raw ~ algae.mean + (1|period) + (1|plot), 
#                     family = gaussian(link = "identity"),
#                     data = algae_means.coral_change.raw)
# summary(coral.algae.glmm.raw)
# 
# # Get confidence interval for model
# confint(coral.algae.glmm.raw)
# 
# # Check residuals
# hist(residuals(coral.algae.glmm.raw))
# car::qqPlot(residuals(coral.algae.glmm.raw))
# 
# # Plot model predictions and create dataframe 
# plot(ggpredict(coral.algae.glmm.raw, terms = ~algae.mean))
# coral.algae.predictions.raw <- as.data.frame(ggpredict(coral.algae.glmm.raw, terms = ~algae.mean))

```

```{r Visualizaton of raw coral change ~ mean algae cover in each period}

# # Algae multiplied by 100 to be expressed in percent cover
# 
# ggplot(coral.algae.predictions.raw, aes(x = x*100, y = predicted))+
#   geom_ribbon(data = coral.algae.predictions.raw,
#               aes(x = x*100, y = predicted, ymin =conf.low, ymax = conf.high),
#               alpha =0.5,
#               fill = "#3B4F8E")+
#   geom_line()+
#   geom_point(data = algae_means.coral_change.raw,
#              aes(x = algae.mean*100, y = delta_coral.raw, shape = treatment),
#              alpha = 0.6,
#              size = 2)+
#   labs(x = "% cover of macroalgae",
#        y = Change~"in"~live~coral~surf.~area~(m^2/yr)) +
#   scale_shape_manual(values = c(19,17),
#                      labels = c("Removal", "Retention"),
#                      name = "Skeletal treatment") +
#   theme_classic(base_size = 14)
```

```{r Stats for raw change in coral ~ raw algae cover}

# Linear mixed effects model of raw change in surface area of live coral ~ macroalgae cover (as a proportion) in a given year; year and plot ID as random effects
coral.algae.glmm.raw.2 <- glmmTMB(delta_coral.raw ~ total_cover.prop + (1|period) + (1|plot), 
                    family = gaussian(link = "identity"),
                    data = algae_totals.coral_change.raw)
summary(coral.algae.glmm.raw.2)

# Get confidence interval for model
confint(coral.algae.glmm.raw.2)

# Check residuals
hist(residuals(coral.algae.glmm.raw.2))
car::qqPlot(residuals(coral.algae.glmm.raw.2))

# Plot model predictions and create dataframe 
plot(ggpredict(coral.algae.glmm.raw.2, terms = ~total_cover.prop))
coral.algae.predictions.raw.2 <- as.data.frame(ggpredict(coral.algae.glmm.raw.2, terms = ~total_cover.prop))

```

```{r Visualizaton of raw coral change ~ mean algae cover in each period}

# Algae multiplied by 100 to be expressed in percent cover

ggplot(coral.algae.predictions.raw.2, aes(x = x*100, y = predicted))+
  geom_ribbon(data = coral.algae.predictions.raw.2,
              aes(x = x*100, y = predicted, ymin =conf.low, ymax = conf.high),
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line()+
  geom_point(data = algae_totals.coral_change.raw,
             aes(x = total_cover.prop*100, y = delta_coral.raw, shape = treatment),
             alpha = 0.6,
             size = 2)+
  labs(x = "% cover of macroalgae",
       y = Change~"in"~live~coral~surf.~area~(m^2/yr)) +
  scale_shape_manual(values = c(19,17),
                     labels = c("Removal", "Retention"),
                     name = "Skeletal treatment") +
  theme_classic(base_size = 14)
```

## Plot sizes
```{r}
# Load dataset for plot areas
plot_areas <- read_csv("data/Disturbance plot areas.csv") %>% 
  clean_names()

# Get plot size ranges for each treatment
plot_areas.removal <- plot_areas %>% 
  filter(treatment == "Removal")
range(plot_areas.removal$plot_area)

plot_areas.retention <- plot_areas %>% 
  filter(treatment == "Retention")
range(plot_areas.retention$plot_area)

# Calculate average plot areas and SE's
plot_areas.summary <- plot_areas %>% 
  group_by(treatment) %>% 
  summarize(mean = mean(plot_area),
            SE = sd(plot_area)/sqrt(10))

# T-test of plot area~treamtent
plot_area.t_test <- t.test(plot_area ~ treatment, data = plot_areas)

```


## Recruit counts
```{r  Load datasets}

recruit_counts <- read_csv("data/Recruit-counts_2023.csv") %>% 
  clean_names() 

```

```{r Recruit counts}
## Total counts
# Sum up recruits found in each treatment and on each substrate 
recruit_counts.summary <- recruit_counts %>%
  #filter(substrate != "Rubble") %>% 
  group_by(treatment, substrate) %>% 
  reframe(total_count = n()) %>% 
  drop_na()

# Visualization of total recruit count ~ treatment and substrate type
names(cal_palettes)

ggplot(recruit_counts.summary, aes(x = treatment, y = total_count, fill = substrate)) +
  geom_col(position = "dodge",
           color = "black") +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Skeleton treatment",
       y = "Total recruit count") +
  scale_fill_manual(values = c("#9BB1BB", "#E4DECE", "#0B4221"),
                    name = "Substrate type") +
  theme_classic(base_size = 14)

## Proportional counts
# calculate proportions of recruits on each substrate type
recruit_counts.summary$proportion <- recruit_counts.summary$total_count / c(rep(44,2),rep(54,2))

# Visualization of total recruit count ~ treatment and substrate type
ggplot(recruit_counts.summary, aes(x = treatment, y = proportion, fill = substrate)) +
  geom_col(position = "dodge",
           color = "black") +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Skeleton treatment",
       y = "Proportion of recruits") +
  scale_fill_manual(values = c("#9BB1BB", "#E4DECE", "#0B4221"),
                    name = "Substrate type") +
  scale_y_continuous(limits = c(0,1),
                     expand = c(0,0)) +
  theme_classic(base_size = 14)

```

```{r Stats for recruit counts ~ substrate and treatment}

recruit_counts.contingency <- recruit_counts.summary %>% 
  filter(substrate != "Rubble") %>% 
  pivot_wider(names_from = "substrate",
              values_from = "total_count") %>% 
  select(-treatment)

recruit.matrix <- as.matrix(recruit_counts.contingency)
colnames(recruit.matrix) <- c("Dead coral", "Reef")
rownames(recruit.matrix) <- c("Removal", "Retention")

recruit.chi_sq <- chisq.test(recruit.matrix)

```

Recruit densities
```{r}

# Calculate recruit densities
recruit_density <- recruit_counts %>% 
  mutate(substrate = case_when(substrate == 'Reef' ~ 'Reef',
                               TRUE ~ 'Dead coral')) %>% 
  group_by(plot_number, treatment, substrate) %>% 
  summarize(total_count = n())

recruit_density <- merge(recruit_density, plot_areas) 

recruit_density <- recruit_density %>%  
  mutate(density = total_count/plot_area)

recruit_density.summary <- recruit_density %>% 
  group_by(treatment, substrate) %>% 
  summarize(mean_density = mean(density),
            se_density = sd(density)/sqrt(n()))

# Visualization
ggplot(recruit_density.summary, aes(x = treatment, y = mean_density, fill = substrate)) +
  geom_col(position = position_dodge(width = 0.9),
           color = "black") +
  geom_errorbar(aes(ymax = mean_density + se_density,
                    ymin = mean_density - se_density),
                position = position_dodge(width = 0.9),
                width = 0) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Skeleton treatment",
       y = Density~of~recruits~(no./m^2)) +
  scale_fill_manual(values = c("#E4DECE", "#0B4221"),
                    name = "Substrate type") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic(base_size = 14)

```

```{r}

# Two-factor ANOVA of recruit density ~ substrate and treatment 
density_aov <- aov(density ~ substrate*treatment, data = recruit_density)
summary(density_aov)

# Posthoc comparisons
emmeans_results <- emmeans(density_aov, ~ substrate * treatment)
pairs(emmeans_results, adjust = "tukey")

emmeans(density_aov, pairwise ~ treatment | substrate, adjust = "tukey")

library(ggplot2)

ggplot(recruit_density, aes(x = substrate, y = density, color = treatment, group = treatment)) +
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0, color = "black") +
  stat_summary(geom = "line", fun = mean, color = "black") +
  stat_summary(geom = "point", fun = mean, size = 4, color = "black") +
  stat_summary(geom = "point", fun = mean, size = 3) +
  scale_color_manual(values = c("#E4DECE", "#0B4221"),
                     name = "Skeleton treatment") +
  theme_minimal() +
  labs(x = "Substrate type",
       y = Density~of~recruits~(no./m^2))

# ANOVA
Anova(glm_model, type = "III") # From the car package

# GLMM method
hist(log(recruit_density$density))

density.glmm <- glmmTMB(log(density) ~ substrate*treatment + (1|plot_number), 
                 data = recruit_density, 
                 family = gaussian)
summary(density.glmm)

density.glmm.2 <- glmmTMB(density ~ substrate*treatment + (1|plot_number), 
                 data = recruit_density, 
                 family = gaussian(link = "log"))
summary(density.glmm.2)

# Posthoc test for comparing substrates within a treatment
emmeans(density.glmm.2, pairwise ~ substrate | treatment, adjust = "tukey")

# Posthoc test for comparing treatments within a substrate
emmeans(density.glmm.2, pairwise ~ treatment | substrate, adjust = "tukey")


# Check residuals
hist(residuals(density.glmm.2))
car::qqPlot(residuals(density.glmm.2))

# Plot model predictions and create dataframe 
plot(ggpredict(density.glmm.2, terms = ~substrate*treatment))
density.predictions <- as.data.frame(ggpredict(density.glmm.2, terms = ~substrate*treatment))

# Model visualization
ggplot(density.predictions, aes(x = x, y = predicted, shape = group, group = group)) +
    # geom_point(data = recruit_density,
    #            aes(x = substrate, y = density, color = treatment, group = treatment),
    #            alpha = 0.6,
    #            position = position_dodge(width = 0.3)) +
    geom_errorbar(aes(ymax = conf.high,
                      ymin = conf.low,
                      group = group),
                  width = 0,
                  color = "black",
                  position = position_dodge(width = 0.3)) +
    geom_line(color = "black",
              position = position_dodge(width = 0.3)) +
    geom_point(size = 4,
               color = "black",
               position = position_dodge(width = 0.3)) +
    geom_point(size = 3,
               position = position_dodge(width = 0.3),
               color = "#CACFD0") +
    scale_shape_manual(values = c(19,17),
                     labels = c("Removal", "Retention"),
                     name = "Skeleton treatment") +
    theme_minimal() +
    labs(x = "Substrate type",
       y = Density~of~recruits~(no./m^2))

# Swap axes
ggplot(density.predictions, aes(x = group, y = predicted, color = x, group = x)) +
    # geom_point(data = recruit_density,
    #            aes(x = substrate, y = density, color = treatment, group = treatment),
    #            alpha = 0.6,
    #            position = position_dodge(width = 0.3)) +
    geom_errorbar(aes(ymax = conf.high,
                      ymin = conf.low,
                      group = x),
                  width = 0,
                  color = "black",
                  position = position_dodge(width = 0.3)) +
  scale_color_manual(values = c("#0B4221", "#E4DECE"),
                     labels = c("Dead coral", "Reef"),
                     name = "Subtrate") +
    geom_line(color = "black",
               position = position_dodge(width = 0.3),
              linetype = 4) +
    geom_point(size = 4,
               color = "black",
               position = position_dodge(width = 0.3)) +
    geom_point(size = 3,
               position = position_dodge(width = 0.3),
               aes(color = x)) +
    theme_minimal() +
    labs(x = "Skelton treatment",
       y = Density~of~recruits~(no./m^2))
  
```



## Supplemental figures
#### Dead coral time series
```{r}
# create dataframe of dead coral cover
dead_coral <- regions_master %>%
  filter(tag_lab_class_name == "Dead coral") %>% 
  group_by(plot, treatment, tag_lab_date) %>% 
  summarize(dead_coral.m_sq = sum(tag_lab_surf_area)*0.0001) %>% 
  mutate(time_point = case_when(tag_lab_date == "8/1/19" ~ 2019,
                                       tag_lab_date == "8/1/20" ~ 2020,
                                       tag_lab_date == "8/1/21" ~ 2021,
                                       tag_lab_date == "8/1/22" ~ 2022,
                                       tag_lab_date == "8/1/23" ~ 2023))

# Dead coral ranges
removal.range <- dead_coral %>% 
  filter(treatment == "Removal")
range(removal.range$dead_coral.m_sq)

retention.range <- dead_coral %>% 
  filter(treatment == "Retention")
range(retention.range$dead_coral.m_sq)

# Create summary of dead coral
dead_coral.summary <- dead_coral %>% 
  group_by(treatment, time_point) %>% 
  summarize(mean_surf.m_sq = mean(dead_coral.m_sq),
            SE_surf.m_sq = sd(dead_coral.m_sq)/sqrt(n()))

# time series of dead coral
ggplot(dead_coral.summary, aes(x = time_point, y = mean_surf.m_sq, shape = treatment)) +
   geom_point(data = dead_coral, aes(x = time_point, y = dead_coral.m_sq),
             size = 1,
             alpha = 0.5) +
  geom_line(aes(group = treatment)) + 
  geom_errorbar(aes(ymax = mean_surf.m_sq + SE_surf.m_sq,
                    ymin = mean_surf.m_sq - SE_surf.m_sq),
                width = 0) +
  geom_point(size = 2.5) +
  geom_point(size = 1.75,
             color = "#CACFD0") +
  labs(x = "Year",
       y = Surface~area~(m^2)~of~dead~coral) +
  scale_shape_manual(values = c(19,17),
                     labels = c("Removal", "Retention"),
                     name = "Skeleton treatment") +
  theme_classic()

# Just dead coral in 2023
dead.2023 <- dead_coral %>% 
  filter(time_point == "2023")

dead.2023.summary <- dead.2023 %>% 
  group_by(treatment) %>% 
  summarize(mean_surf.m_sq = mean(dead_coral.m_sq),
            SE_surf.m_sq = sd(dead_coral.m_sq)/sqrt(n()))

# t-test of dead coral cover ~ treatment in 2023
dead.t_test <- t.test(dead_coral.m_sq ~ treatment, data = dead.2023)

# time series of dead coral in 2023 only
ggplot(dead.2023.summary, aes(x = treatment, y = mean_surf.m_sq, shape = treatment)) +
   geom_point(data = dead.2023, 
             aes(x = treatment, y = dead_coral.m_sq),
             size = 1.5,
             alpha = 0.5) +
  geom_errorbar(aes(ymax = mean_surf.m_sq + SE_surf.m_sq,
                    ymin = mean_surf.m_sq - SE_surf.m_sq),
                width = 0) +
  geom_point(size = 3) +
  geom_point(size = 2,
             color = "#CACFD0") +
  labs(x = "Skeleton treatment",
       y = Surface~area~(m^2)~of~dead~coral) +
  scale_shape_manual(values = c(19,17),
                     #labels = c("Skeleton removal", "Skeletal retention"),
                     name = "") +
  scale_y_continuous(limits = c(0,5),
                     expand = c(0,0)) +
  coord_fixed(ratio = 1) +
  theme_classic(base_size = 14)
```

#### Time series of live coral raw values
```{r}

# Calculate means and SE's of raw surface area of live coral for each plot at each timepoint 
live_coral.summary <- live_coral %>% 
  group_by(treatment, time_point) %>% 
  summarize(mean_surf = mean(live_coral.total_surf),
            SE_surf = sd(live_coral.total_surf)/sqrt(n()),
            mean_surf.m_sq = mean(m_sq),
            SE_surf.m_sq = sd(m_sq)/sqrt(n()))

# Time series of raw values (m^2)
ggplot(live_coral.summary, aes(x = time_point, y = mean_surf.m_sq, shape = treatment)) +
  geom_line(aes(group = treatment)) + 
  geom_errorbar(aes(ymax = mean_surf.m_sq + SE_surf.m_sq,
                    ymin = mean_surf.m_sq - SE_surf.m_sq),
                width = 0) +
   geom_point(data = live_coral, aes(x = time_point, y = m_sq),
             size = 1,
             alpha = 0.5) +
  geom_point(size = 2.5) +
  geom_point(size = 2,
             color = "#CACFD0") +
  labs(x = "Year",
        y = Surface~area~(m^2)~of~live~coral) +
  scale_shape_manual(values = c(19,17),
                     labels = c("Removal", "Retention"),
                     name = "Skeleton treatment") +
  theme_classic()
```

MCR coral time series for supplement: coral taxa by year
```{r}
mcr <- read_csv("Data/MCR_coral data.csv") %>% 
   clean_names()

mcr.2023 <- read_csv("Data/2023_LTER_1_10m_Coral_Analysis.csv") %>% 
   clean_names()

mcr_coral <- mcr %>% 
  mutate(site = str_sub(location, 0, 6),
         year = str_sub(date, 0, 4)) %>% 
  filter(str_detect(location, "Outer 10 m"),
         site == "LTER 1",
         year %in% c(2018, 2019, 2020, 2021, 2022, 2023), 
         taxonomy_substrate_or_functional_group != "Non-coralline Crustose Algae",
         taxonomy_substrate_or_functional_group != "CTB",
         taxonomy_substrate_or_functional_group != "Macroalgae",
         taxonomy_substrate_or_functional_group != "Millepora",
         taxonomy_substrate_or_functional_group != "Sand",
         taxonomy_substrate_or_functional_group != "Soft Coral",
         taxonomy_substrate_or_functional_group != "Unknown or Other",
         taxonomy_substrate_or_functional_group != "Turf") %>% 
  rename(Taxa = taxonomy_substrate_or_functional_group) %>% 
  mutate(percent_cover = as.numeric(percent_cover)) %>% 
  drop_na(percent_cover)%>% 
  select(-c(5:7))

mcr_coral.2023 <- mcr.2023 %>% 
  mutate(site = str_sub(location, 0, 6),
         year = str_sub(date, 0, 4)) %>% 
  filter(str_detect(location, "Outer 10 m"),
         site == "LTER 1",
         year %in% c(2018, 2019, 2020, 2021, 2022, 2023), 
         taxonomy != "Crustose Coralline Algae / Bare Space",
         taxonomy != "Non-coralline Crustose Algae",
         taxonomy != "Macroalgae",
         taxonomy != "Millepora",
         taxonomy != "Sand",
         taxonomy != "Soft Coral",
         taxonomy != "Unknown or Other",
         taxonomy != "Turf") %>% 
  rename(Taxa = taxonomy) %>% 
  mutate(percent_cover = as.numeric(percent_cover)) %>% 
  drop_na(percent_cover) %>% 
  select(-c(transect, quadrat))

mcr_coral <- rbind(mcr_coral, mcr_coral.2023)

# Total up percent cover of different taxa to get mean cover of each coral taxa
mcr_coral.totals <- mcr_coral %>% 
  group_by(year, Taxa) %>% 
  summarize(percent_cover.mean = mean(percent_cover),
            percent_cover.se = sd(percent_cover)/sqrt(n()))

# Mean coral cover by taxa from 2018-2023
ggplot(mcr_coral.totals, aes(x = year, y = percent_cover.mean, fill = Taxa)) +
  geom_col(position = position_dodge(width = 0.8), color = "black") +
  geom_errorbar(aes(ymax = percent_cover.mean + percent_cover.se,
                    ymin = percent_cover.mean - percent_cover.se),
                position = position_dodge(width = 0.8),
                width = 0.2) +
  labs(x = "Year",
       y = "Percent cover",
       fill = "Coral taxa") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic(base_size = 14) +
  theme(legend.position = c(0.95, 0.95),
        legend.justification = c(1, 1),  
        legend.background = element_rect(fill = "white", color = "black"))  

```

